---
title: "Metafor for LLM Numerical Data Extraction (Case Study)"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We manually extract the relevant information and then plug the values into Metafor's escalc function.
Then, we fit random-effects model (fixed effects) and create a forest plot for visualization.

## Metafor's escalc

Reference: https://wviechtb.github.io/metafor/reference/escalc.html

Function to calculate various effect sizes or outcome measures (and the corresponding sampling variances) that are commonly used in meta-analyses.

```
escalc(measure, ai, bi, ci, di, n1i, n2i, x1i, x2i, t1i, t2i,
       m1i, m2i, sd1i, sd2i, xi, mi, ri, ti, fi, pi, sdi, r2i, ni, yi, vi, sei,
       data, slab, subset, include,
       add=1/2, to="only0", drop00=FALSE, vtype="LS",
       var.names=c("yi","vi"), add.measure=FALSE,
       append=TRUE, replace=TRUE, digits, ...)
```

## Metafor's rma

Reference: https://wviechtb.github.io/metafor/reference/rma.uni.html

Function to fit meta-analytic equal-, fixed-, and random-effects models and (mixed-effects) meta-regression models using a linear (mixed-effects) model framework. In this work, we do a standard fixed-effects models.

```
rma(yi, vi, sei, weights, ai, bi, ci, di, n1i, n2i, x1i, x2i, t1i, t2i,
        m1i, m2i, sd1i, sd2i, xi, mi, ri, ti, fi, pi, sdi, r2i, ni, mods, scale,
        measure="GEN", intercept=TRUE, data, slab, subset,
        add=1/2, to="only0", drop00=FALSE, vtype="LS",
        method="REML", weighted=TRUE, test="z",
        level=95, btt, att, tau2, verbose=FALSE, digits, control, ...)
```

## Metafor's forest

Reference: https://wviechtb.github.io/metafor/reference/forest.html

Function to create forest plots.

```
forest(x, vi, sei, ci.lb, ci.ub,
       annotate=TRUE, showweights=FALSE, header=FALSE,
       xlim, alim, olim, ylim, at, steps=5,
       level=95, refline=0, digits=2L, width,
       xlab, slab, ilab, ilab.xpos, ilab.pos,
       order, subset, transf, atransf, targs, rows,
       efac=1, pch, psize, plim=c(0.5,1.5), col,
       shade, colshade, lty, fonts, cex, cex.lab, cex.axis, ...)
```

## Import package

Import relevant package:

```{r packages, message=FALSE}
# install.packages("metafor")
# load metafor package
library(metafor)
```

## Case Study

We calculate the log odds ratios, fit meta-analytic fixed-effects model, and create the forest plots for cochrane reference meta anlayses and outputs from two LLMs (GPT-4 and Mistral 7B Instruct).

### Cochrane

```{r forest_cochrane_reference}
### Get the data
dat <- read.csv("files/cochrane_binary_outcomes.csv")
 
### calculate log odds ratios and corresponding sampling variances (and use
### the 'slab' argument to store study labels as part of the data frame)
dat <- escalc(measure="OR", ai=ai, n1i=n1i, ci=ci, n2i=n2i, data=dat,
              slab=paste(author, year, sep=", "), drop00=TRUE)
dat
 
### fit random-effects model (fixed effects)
res <- rma(yi, vi, data=dat, method = "FE")
res
 
### estimated average odds ratio (and 95% CI/PI)
pred <- predict(res, transf=exp, digits=2)
pred
 
############################################################################
 
### need the rounded estimate and CI bounds further below
pred <- fmtx(c(pred$pred, pred$ci.lb, pred$ci.ub), digits=2)
 
### total number of studies
k <- nrow(dat)

### get the weights and format them as will be used in the forest plot
weights <- paste0(fmtx(weights(res), digits=1), "%")
weights[weights == "NA%"] <- ""
 
### adjust the margins
par(mar=c(13.8,0,1.3,0), mgp=c(2,0.2,0), tcl=-0.2)

pdf(width=8.8,height=3.2,file="forest_plots/cochrane_forest_plot.pdf")
 
### forest plot with extra annotations
sav <- forest(res, atransf=exp, at=log(c(0.01, 0.10, 1, 10, 100)), xlim=c(-30,11),
       xlab="", efac=c(0,4), textpos=c(-30,-4.7), lty=c(1,1,0), refline=NA,
       ilab=cbind(ai, n1i, ci, n2i, weights),
       ilab.xpos=c(-20.6,-18.6,-16.1,-14.1,-10.8), ilab.pos=2,
       cex=0.78, header=c("Study"), mlab="")

### add horizontal line at the top
segments(sav$xlim[1], k+1, sav$xlim[2]-5, k+1, lwd=0.8)
 
### add vertical reference line at 0
segments(0, -2, 0, k+1, lwd=0.8)
 
### now we add a bunch of text; since some of the text falls outside of the
### plot region, we set xpd=NA so nothing gets clipped
par(xpd=NA)
 
### adjust cex as used in the forest plot and use a bold font
par(cex=sav$cex, font=2)

text(sav$ilab.xpos, k+2, pos=2, c("Events","Total","Events","Total", "Weight"))
text(c(-19.1,-15.1), k+3, pos=2, c("Remdesivir","Control"))
text(0, k+3, "Odds ratio, 95% CI")
 
### use a non-bold font for the rest of the text
par(cex=sav$cex, font=1)

text(c(sav$xlim[1],sav$ilab.xpos[c(2,4,5)]), -1, pos=c(4,2,2,2,2),
     c("Total (95% CI)", sum(dat$n1i), sum(dat$n2i), "100.0%"))
 
### add 'Favors remdesivir'/'Favors control' text below the x-axis
text(log(c(0.01, 100)), -3, c("Favors remdesivir","Favors control"), pos=c(4,2), offset=-1)
 
### add text for total events
text(sav$ilab.xpos[c(1,3)], -1, c(sum(dat$ai),sum(dat$ci)), pos=2)

dev.off
```

### GPT-4

```{r gpt4_binary_outcomes}
### Get the data
dat <- read.csv("files/gpt4_binary_outcomes.csv")
dat

### calculate log odds ratios and corresponding sampling variances (and use
### the 'slab' argument to store study labels as part of the data frame)
dat <- escalc(measure="OR", ai=ai, n1i=n1i, ci=ci, n2i=n2i, data=dat,
              slab=paste(author, year, sep=", "), drop00=TRUE)
dat
 
### fit random-effects model (fixed effects)
res <- rma(yi, vi, data=dat, method = "FE")
res
 
### estimated average odds ratio (and 95% CI/PI)
pred <- predict(res, transf=exp, digits=2)
pred
 
############################################################################
 
### need the rounded estimate and CI bounds further below
pred <- fmtx(c(pred$pred, pred$ci.lb, pred$ci.ub), digits=2)
 
### total number of studies
k <- nrow(dat)

### get the weights and format them as will be used in the forest plot
weights <- paste0(fmtx(weights(res), digits=1), "%")
weights[weights == "NA%"] <- ""
 
### adjust the margins
par(mar=c(13.8,0,1.3,0), mgp=c(2,0.2,0), tcl=-0.2)

pdf(width=8.8,height=3.2,file="forest_plots/gpt4_forest_plot.pdf")
 
### forest plot with extra annotations
sav <- forest(res, atransf=exp, at=log(c(0.01, 0.10, 1, 10, 100)), xlim=c(-30,11),
       xlab="", efac=c(0,4), textpos=c(-30,-4.7), lty=c(1,1,0), refline=NA,
       ilab=cbind(ai, n1i, ci, n2i, weights),
       ilab.xpos=c(-20.6,-18.6,-16.1,-14.1,-10.8), ilab.pos=2,
       cex=0.78, header=c("Study"), mlab="")

### add horizontal line at the top
segments(sav$xlim[1], k+1, sav$xlim[2]-5, k+1, lwd=0.8)
 
### add vertical reference line at 0
segments(0, -2, 0, k+1, lwd=0.8)
 
### now we add a bunch of text; since some of the text falls outside of the
### plot region, we set xpd=NA so nothing gets clipped
par(xpd=NA)
 
### adjust cex as used in the forest plot and use a bold font
par(cex=sav$cex, font=2)

text(sav$ilab.xpos, k+2, pos=2, c("Events","Total","Events","Total", "Weight"))
text(c(-19.1,-15.1), k+3, pos=2, c("Remdesivir","Control"))
text(0, k+3, "Odds ratio, 95% CI")
 
### use a non-bold font for the rest of the text
par(cex=sav$cex, font=1)

text(c(sav$xlim[1],sav$ilab.xpos[c(2,4,5)]), -1, pos=c(4,2,2,2,2),
     c("Total (95% CI)", sum(dat$n1i), sum(dat$n2i), "100.0%"))
 
### add 'Favors remdesivir'/'Favors control' text below the x-axis
text(log(c(0.01, 100)), -3, c("Favors remdesivir","Favors control"), pos=c(4,2), offset=-1)
 
### add text for total events
text(sav$ilab.xpos[c(1,3)], -1, c(sum(dat$ai),sum(dat$ci)), pos=2)

dev.off
```

### Mistral Instruct 7B

```{r mistral_binary_outcomes}
### Get the data
dat <- read.csv("files/mistral7B_binary_outcomes.csv")
dat

### calculate log odds ratios and corresponding sampling variances (and use
### the 'slab' argument to store study labels as part of the data frame)
dat <- escalc(measure="OR", ai=ai, n1i=n1i, ci=ci, n2i=n2i, data=dat,
              slab=paste(author, year, sep=", "), drop00=TRUE)
dat
 
### fit random-effects model (fixed effects)
res <- rma(yi, vi, data=dat, method = "FE")
res
 
### estimated average odds ratio (and 95% CI/PI)
pred <- predict(res, transf=exp, digits=2)
pred
 
############################################################################
 
### need the rounded estimate and CI bounds further below
pred <- fmtx(c(pred$pred, pred$ci.lb, pred$ci.ub), digits=2)
 
### total number of studies
k <- nrow(dat)

### get the weights and format them as will be used in the forest plot
weights <- paste0(fmtx(weights(res), digits=1), "%")
weights[weights == "NA%"] <- ""
 
### adjust the margins
par(mar=c(13.8,0,1.3,0), mgp=c(2,0.2,0), tcl=-0.2)

pdf(width=8.8,height=3.2,file="forest_plots/mistral_forest_plot.pdf")
 
### forest plot with extra annotations
sav <- forest(res, atransf=exp, at=log(c(0.01, 0.10, 1, 10, 100)), xlim=c(-30,11),
       xlab="", efac=c(0,4), textpos=c(-30,-4.7), lty=c(1,1,0), refline=NA,
       ilab=cbind(ai, n1i, ci, n2i, weights),
       ilab.xpos=c(-20.6,-18.6,-16.1,-14.1,-10.8), ilab.pos=2,
       cex=0.78, header=c("Study"), mlab="")

### add horizontal line at the top
segments(sav$xlim[1], k+1, sav$xlim[2]-5, k+1, lwd=0.8)
 
### add vertical reference line at 0
segments(0, -2, 0, k+1, lwd=0.8)
 
### now we add a bunch of text; since some of the text falls outside of the
### plot region, we set xpd=NA so nothing gets clipped
par(xpd=NA)
 
### adjust cex as used in the forest plot and use a bold font
par(cex=sav$cex, font=2)

text(sav$ilab.xpos, k+2, pos=2, c("Events","Total","Events","Total", "Weight"))
text(c(-19.1,-15.1), k+3, pos=2, c("Remdesivir","Control"))
text(0, k+3, "Odds ratio, 95% CI")
 
### use a non-bold font for the rest of the text
par(cex=sav$cex, font=1)

text(c(sav$xlim[1],sav$ilab.xpos[c(2,4,5)]), -1, pos=c(4,2,2,2,2),
     c("Total (95% CI)", sum(dat$n1i), sum(dat$n2i), "100.0%"))
 
### add 'Favors remdesivir'/'Favors control' text below the x-axis
text(log(c(0.01, 100)), -3, c("Favors remdesivir","Favors control"), pos=c(4,2), offset=-1)
 
### add text for total events
text(sav$ilab.xpos[c(1,3)], -1, c(sum(dat$ai),sum(dat$ci)), pos=2)

dev.off
```
